---
title: "R Notebook"
output: html_notebook
---
```{r}
library(car)
```


# Loading the cleaned dataset
```{r}
file <- 'life_expectancy_final_data.csv'
dat <- read.csv('/Users/admin/Downloads/life_expectancy_final_data.csv')
```
```{r}
summary(dat)
```

```{r}
head(dat)
```
# Model Selection 

#Backward Elimination
```{r}
out.full=lm(life_expectancy~.- X- status - country - year,dat)
out.backward=step(out.full,scope=list(lower=~1,upper=full),direction="backward",trace=FALSE,k=log(n))
out.backward$coefficients
```
```{r}
temp=lm(out.full)
summary(temp)
AIC(temp)
BIC(temp)
vif(temp)
```

```{r}
vif(out.backward)
```

```{r}
out.backward_1=lm(life_expectancy~adult_mortality+ alcohol + bmi + polio+ total_expenditure + diphtheria + hiv.aids + gdp + thinness__1_19_years + income_composition_of_resources, data=dat )
out.backward_1$coefficients
```

```{r}
summary(out.backward_1)
AIC(out.backward_1)
BIC(out.backward_1)
```

#Finding multicollinearity
```{r}
vif(out.backward_1)
```


# Forward Selection
```{r}
n = nrow(dat)
out.null=lm(life_expectancy~1, data=dat)
full=formula(lm(life_expectancy~. - X - status - country - year,dat))
out.forward=step(out.null,scope=list(lower=~1,upper=full),k=log(n),direction="forward",trace=FALSE)
out.forward$coefficients
```
```{r}
vif(out.forward)
```


```{r}
summary(out.forward)
AIC(out.forward)
BIC(out.forward)
```

# We conclude that the forward selection is better over bckward elimination because it hs a higher adjusted R^2, lower AIC and BIC

```{r}
plot(fitted(out.forward),residuals(out.forward))
abline(0,0)
```

#Normality plots
```{r}
qqnorm(residuals(out.forward))
qqline(residuals(out.forward))
```

#From the normality plot, we observe that the distribution is a bit light tailed. So,we try to perform box cox transform to make the distribution close to normal

```{r}
plot(density(residuals(out.forward)))
```

#Boxcox Transformation
```{r}
model_3= lm((((life_expectancy)**(-0.802))-1)/(-0.802) ~ schooling + adult_mortality + hiv.aids + diphtheria + bmi + income_composition_of_resources + percentage_expenditure + polio + thinness__1_19_years + measles + alcohol,data=dat)
summary(model_3)

```

```{r}
qqnorm(residuals(model_3))
qqline(residuals(model_3))
```
#From the normality plots of the transformed model, we observe that the distribution is still a little light tailed but its better compared to the previous model. Hence, this model is better
```{r}
tail(dat)
```

# Finding Outliers
```{r}
student <- rstudent(model_3)
jackkniferes <- student*(2925/(2926-student^2))^0.5
head(jackkniferes[order(abs(student),decreasing=T)],15)

```

```{r}
qt(0.025/2938,2925)

```

#Dropping the outliers
```{r}
df_drop1 <- dat[-c(63,64,1128,2306,2308,2310,2312,2313,2309,2307,434),]
```

#After dropping the outliers we fit the final model 
```{r}
model_final= lm((((life_expectancy)**(-0.802))-1)/(-0.802) ~ schooling + adult_mortality + hiv.aids + diphtheria + bmi + income_composition_of_resources + percentage_expenditure + polio + thinness__1_19_years + measles + alcohol,data=df_drop1)
summary(model_final)
```
#We observe the adjusted R^2 is 0.8124 and all the predictors p value is way less than the 5% significance level(0.05). Hence, this is model is a good fit


#Now lets verify if there are any more outliers
```{r}
student <- rstudent(model_final)
jackkniferes <- student*(2914/(2915-student^2))^0.5
head(jackkniferes[order(abs(student),decreasing=T)])
```

```{r}
qt(0.025/2927,2914)

```
#Conclusion -There are no mpre outliers

#Plotting the normality for the final model
```{r}
qqnorm(residuals(model_final))
qqline(residuals(model_final))
```
#We see this is a very good fit almost close to the normal line

#Plotting the residuls
```{r}
plot(fitted(model_final),residuals(model_final))
abline(0,0)
```

#Plotting the density
```{r}
plot(density(residuals(model_final)))

```
#We see that the distribution is almost normal. Hence, we conclude that this model is a very good fit


#Influential Points
```{r}
library('faraway')
states<-row.names(dat)
cook<-cooks.distance(out.forward)
halfnorm(cook,6,labs=states,ylab="Cook’s distances")
```
# Conclusion


#Hypothesis testing
```{r}

#Test 1
# We realize that none of the immunization factors seem to have correlation on life expectancy but we want to know which of the immunizations factors have a higher impact on life expectancy.

# Test diphteria > polio
#H0 : diphteria = polio
#H1 : diphteria < polio
linearHypothesis(model_final, c('diphtheria - polio = 0'))
```

# Conclusion - As the p value is greter than 0.05, we accept the null hypothesis and conclude that the immunization of diphtheria has more effect on life expectancy than the immunization of polio.


```{r}
#Test 2
# Which of the diseases has more impct
#H0 : measles = hiv.aids
#H1 : measles < hiv.aids
linearHypothesis(model_final, c('measles - hiv.aids = 0'))
```


#Conslusion - As the p value is less than 0.05, we reject the null hypothesis and conclude that hiv has more impact on life expectancy than measles

```{r}
# Test 3
# Is alcohol even a good factor to predict life expectancy
#H0 : alcohol = 0
#H1 : alcohol != 0
linearHypothesis(model_final, c('alcohol = 0'))
```

#Conclusion -As p value is less than 0.05, we reject the null hypothesis and conclude that alcohol has some impact on life expectancy.
#But at 99.9% significance level, alcohol does not have an impact on life expectancy

